---
title: ripeds
output: md_document
---

```{r, include = FALSE}
options(width = 100)
```
# Download and Manipulate Integrated Postsecondary Education Data System (IPEDS) Data

[![R-CMD-check](https://github.com/nehgov/ripeds/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nehgov/ripeds/actions/workflows/R-CMD-check.yaml)
[![GitHub
release](https://img.shields.io/github/release/nehgov/ripeds.svg)](https://github.com/nehgov/ripeds)
![R CMD check status](https://github.com/nehgov/ripeds/actions)
![beta](https://img.shields.io/badge/beta-blue)



The `ripeds` R package provides a series of piped functions to facilitate
downloading Department of Education Integrated Postsecondary Education Data
System (IPEDS) complete data files.

All command pipes must start with `ipeds_init()`, end with `ipeds_get()`, and be
linked with the base pipe, `|>`, or magrittr pipe function, `%>%`. Internal
commands, `ipeds_select()`, `ipeds_filter()`, and `ipeds_year()` come in any
order in the pipe chain. Only `ipeds_select()` is required.

# Installation

```r
devtools::install_github("nehgov/ripeds")
```

# Example

```{r}
## perform data pull
df <- ipeds_init() |>
  ipeds_select(instnm, pt_ug) |>
  ipeds_filter(stabbr == "KY") |>
  ipeds_year(2020:2021) |>
  ipeds_get()

## show
df |> head()
```

# Disclaimer

This software package is licensed under the CC0 license. It is offered as-is and
makes no representations or warranties of any kind concerning the Work, express,
implied, statutory or otherwise, including without limitation warranties of
title, merchantability, fitness for a particular purpose, non infringement, or
the absence of latent or other defects, accuracy, or the present or absence of
errors, whether or not discoverable, all to the greatest extent permissible
under applicable law.
