---
title: ripeds
output: md_document
---

```{r, include = FALSE}
options(width = 100)
```
# Download and Manipulate Integrated Postsecondary Education Data System (IPEDS) Data

[![R-CMD-check](https://github.com/nehgov/ripeds/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nehgov/ripeds/actions/workflows/R-CMD-check.yaml)
![GitHub Release](https://img.shields.io/github/v/release/nehgov/ripeds)
![Static Badge](https://img.shields.io/badge/release-beta-blue)

**THIS SOFTWARE PACKAGE IS IN BETA RELEASE AND MAY CONTAIN SOME BUGS. 
IN ADDITION, SOME FUNCTIONALITY MAY CHANGE. PLEASE SEE DISCLAIMER BELOW 
AND USE ACCORDINGLY.**

The `ripeds` R package provides a series of piped functions to facilitate
downloading Department of Education Integrated Postsecondary Education Data
System (IPEDS) complete data files.

All command pipes must start with `ipeds_init()`, end with `ipeds_get()`, and be
linked with the base pipe, `|>`, or magrittr pipe function, `%>%`. Internal
commands, `ipeds_select()`, `ipeds_filter()`, and `ipeds_year()` come in any
order in the pipe chain. Only `ipeds_select()` is required.

This software package was developed by the Office of Data and Evaluation of the
National Endowment for the Humanities.

# Installation

```r
devtools::install_github("nehgov/ripeds")
```

# Example

```{r}
library(ripeds)

## perform data pull
df <- ipeds_init() |>
  ipeds_select(instnm, pt_ug) |>
  ipeds_filter(stabbr == "KY") |>
  ipeds_year(2020:2021) |>
  ipeds_get()

## show
df |> head()
```
# Help and vignettes

See the [`ripeds`](https://nehgov.github.io/ripeds) site for more help as well
as vignettes on explaining package functionality.

# Disclaimer

The software package is licensed under the CC0 license. It is provided "as is"
without any warranty of any kind, either expressed, implied, or statutory,
including, but not limited to, any warranty that the subject software will
conform to specifications, any implied warranties of merchantability, fitness
for a particular purpose, or freedom from infringement, any warranty that the
subject software will be error free, or any warranty that documentation, if
provided, will conform to the subject software. This agreement does not, in any
manner, constitute an endorsement by NEH or any prior recipient of any results,
resulting designs, hardware, software products or any other applications
resulting from use of the subject software. Further, NEH disclaims all
warranties and liabilities regarding third-party software, if present in the
original software, and distributes it "as is."

The recipient agrees to waive any and all claims against the United States
government, its contractors and subcontractors, as well as any prior recipient.
If recipient's use of the subject software results in any liabilities, demands,
damages, expenses or losses arising from such use, including any damages from
products based on, or resulting from, recipient's use of the subject software,
recipient shall indemnify and hold harmless the United States government, its
contractors and subcontractors, as well as any prior recipient, to the extent
permitted by law. Recipient's sole remedy for any such matter shall be the
immediate, unilateral termination of this agreement.
