---
title: "Batch download"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Batch download}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Depending on your analysis, you may wish to download a number of data files to
your hard disk. This can save time when working on large scale projects. In
addition, downloading and saving the raw data files used in your analysis can
support future reproduction. Should files on the server be revised or modified,
you will have a copy of the files as they existed at the time of your original
analysis.

You can also download the dictionary files that accompany each complete data
file. While `iped_dict()` provides useful information, the full data files may
contain more detailed definitions and distributional information (_e.g._,
frequencies, means, _etc._).


```{r setup}
library(ripeds)
```

## `ipeds_download_to_disk()`

To download either complete data files or the accompanying dictionary files, use
the `ipeds_download_to_disk()` function.

There are two basic ways to download the files:

1. Using file names directly
1. Using results from the helper function `ipeds_file_table()`
   
### Using file names

If you know the file names you need, you can use them directly. Be sure to give
a local directory for the download. If the directory doesn't exist, R will try
to create it, unless you set `create_dir = FALSE`.

```r
## single file to current working directory
ipeds_download_to_disk(to_dir = ".", files = "IC2022")
```

Note that you do not need to specify the file ending, only the file stub name.
Whichever files you choose, you will end up with a zip file
(<file\_stub\_name.zip>: `IC2022.zip`) that contains:

1. The CSV file (typically lowercase): `ic2022.csv`
1. A revised version of the file, if one exists, ending in `*_rv`: `ic2022_rv.csv` 

You can pass an object with vector of names.

```r
## multiple files stored in object
files_to_get <- c("IC2020", "IC2021", "IC2022")
ipeds_download_to_disk(to_dir = ".", files = files_to_get)
```
The zipped version of each of these files will be downloaded to the `to_dir`.

### Using results from helper function `ipeds_file_table()`

If you don't know the specific file names or simply want to download a larger
number of files without listing them all, you can use results from
`ipeds_file_table()`.

For example, if you want to download all files associated with 2020, you can
filter the results of `ipeds_file_table()` to that year:

```r
## save results from ipeds_file_table() to object
dict <- ipeds_file_table()
```
```r
## pull all files from 2020
ipeds_download_to_disk(".", use_ipeds_dict = dict[dict$year == 2020,])
```

Note that for files with academic year naming conventions (_e.g._, 9899), the
fall semester is the year associated with the file: 2020 will select 20-21 files.

You can also filter on other values. Broadly, IPEDS survey components fall into
a few categories. You can filter to return all **Institutional Characteristics**
survey files:

```r
## download all surveys under the Institutional Characteristics category
ipeds_download_to_disk(".", use_ipeds_dict = dict[dict$survey == "Institutional Characteristics",])
```

You can filter by key words in the survey titles. For example, if you
want to save all files related to directory information:

```r
## download using regular expressions to grab files with "Directory" in title
ipeds_download_to_disk(".", use_ipeds_dict = dict[grepl("Directory", dict$title),])
```

Finally, you can download the entirety of IPEDS if you use `ipeds_file_table()` with no filter:

```r
## download all files
ipeds_download_to_disk(".", use_ipeds_dict = dict)
```

Note, however, that IPEDS is very large. It both will take a while to download
and take up substantial disk space (> 1GB as of January 2025). 

## Dictionary files

Download accompanying dictionary files by setting `type = "dictionary"`:
  
```r
## get dictionary file
ipeds_download_to_disk(".", "IC2020", type = "dictionary")
```

These files will be named using the stub name plus `*_Dict` ending and also
stored inside a zip directory: _e.g._, `IC2020_Dict.zip`.

## Overwrite

By default, `ipeds_download_to_disk()` won't overwrite existing files. This is
to save time and bandwidth. If you wish to download all input files, set `overwrite = TRUE`:

```r
## overwrite
files_to_get <- c("IC2020", "IC2021", "IC2022")
ipeds_download_to_disk(to_dir = ".", files = files_to_get, overwrite = TRUE)
```
