---
title: "Non-standard evaluation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Non-standard evaluation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  % VignetteDepends{dplyr,purrr,tibble}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

By default `ripeds` allows for non-standard evaluation (NSE) in the functions.
This default setting will feel normal to those accustomed to working with
packages in the [Tidyverse](https://www.tidyverse.org). Users can also turn off
NSE with `ipeds_init(use_nse = FALSE)`.

```{r setup}
library("ripeds")
```

# Selecting variables

With NSE, variables can be separated by commas without quotation.

**NSE**
```{r}
df <- ipeds_init() |>
  ipeds_select(instnm, stabbr) |>
  ipeds_year(2023) |>
  ipeds_get()
```

```{r}
df |> head()
```

Using standard evaluation, quotation and concatenation `c()` is required.

**Standard**
```{r}
df <- ipeds_init(use_nse = FALSE) |>
  ipeds_select(c("instnm", "stabbr")) |>
  ipeds_year(2023) |>
  ipeds_get()
```

```{r}
df |> head()
```

Note that with standard evaluation, users can store variables names in an object
and pass the object to `ipeds_select()`. This can be useful with large variable
requests or dynamic requests.

```{r}
## using object
vars_to_get <- c("instnm", "stabbr")
df <- ipeds_init(use_nse = FALSE) |>
  ipeds_select(vars_to_get) |>
  ipeds_year(2023) |>
  ipeds_get()

df |> head()
```

```{r}
## dynamically
df_list <- purrr::map(c("instnm", "stabbr", "control"),
                      ~ ipeds_init(use_nse = FALSE) |>
                        ipeds_select(.x) |>
                        ipeds_year(2023) |>
                        ipeds_get())
```

```{r}
lapply(df_list, head)
```

# Filters

Filters can also be applied using either NSE or standard evaluation. With NSE,
multiple filters are separated with commas, which are treated as **AND** (`&`).

**NSE**
```{r}
df <- ipeds_init() |>
  ipeds_select(instnm, stabbr) |>
  ipeds_year(2023) |>
  ipeds_filter(stabbr == "KY", control == 1) |> # stabbr == "KY" & control == 1
  ipeds_get()
```

```{r}
df |> head()
```

```{r}
df |> dplyr::distinct(stabbr, control)
```

If you want a more complex filter, you will likely need to use standard
evaluation.

**Standard**
```{r}
df <- ipeds_init(use_nse = FALSE) |>
  ipeds_select(c("instnm", "stabbr")) |>
  ipeds_year(2023) |>
  ipeds_filter("stabbr == 'KY' & (control == 1 | control == 2)") |> 
  ipeds_get()
```

```{r}
df |> head()
```

```{r}
df |> dplyr::distinct(stabbr, control)
```

Note that with `ipeds_init(use_nse = FALSE)` you will have to use standard
evaluation throughout the call chain. For more information about how filters
work, see the vignette on [filter behavior](./filter_behavior.html).
