---
title: "Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  % VignetteDepends{dplyr,tibble}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup

```{r setup}
#| include: FALSE
libs <- c("dplyr", "tibble", "ripeds")
sapply(libs, require, character.only = TRUE)
```

# Example 1: institution locations

For this first request, we'll search for institution coordinates. Since we don't
know the exact variable names, we can use `ipeds_dict()`. By default,
`ipeds_dict()` will search variable names, descriptions, and filenames. We'll
start with **latitude**:

## Search dictionary
```{r}
ipeds_dict("latitude")
```

The dictionary output shows two variables, _pclatitude_ and _latitude_. Both
have similar descriptions, but _latitude_ has a longer stretch of years
available. Now we'll search for **longitude**:

```{r}
ipeds_dict("longitude")
```

Similar output, but we note that **longitude** is represented by a variable that
drops the final **e**: _longitud_. We have our variables.

## Pull most recent year

To begin, we pull only the most recent year, which we can see from the output of
the data dictionary under `../FILES`: 2023. We'll run the default chain and add
a call to `dplyr::as_tibble()` at the end to covert the data.frame output to a
tibble for nicer viewing:

```{r}
df <- ipeds_init() |>
  ipeds_select(latitude, longitud) |>
  ipeds_year(2023) |> 
  ipeds_get() |>
  as_tibble()
```
```{r}
df
```

The output contains the requested variables, plus the unique institutional
_unitid_, and year associated with the request.

## Pull past 10 years

Modifying the request slightly, we can request the past 10 years of data. Most
institutions don't move, but some do and that might be interesting!

```{r}
df <- ipeds_init() |>
  ipeds_select(latitude, longitud) |>
  ipeds_year(2014:2023) |> 
  ipeds_get() |>
  as_tibble() |>
  arrange(unitid, year)
```
```{r}
df
```

## Add other useful variables as well as a filter

In the next step, we can add more variables, including _instnm_ (institution
name), and _sector_, which gives the level (sub two-year, two-year, and
four-year) by control (public, private not-for-profit, and private for-profit).
We can also limit to institutions in Kentucky by using the filter `stabbr ==
"KY"`. Because neither the additional variables nor the filtering variables come
from different files than those already downloaded and in memory, this request
should be faster.

```{r}
df <- ipeds_init() |>
  ipeds_select(instnm, sector, latitude, longitud) |>
  ipeds_year(2014:2023) |>
  ipeds_filter(stabbr == "KY") |> 
  ipeds_get() |>
  as_tibble()
```
```{r}
df
```

# Example 2: exclusively distance undergraduate students

For this example, we'll search for the number of exclusively distance education
undergraduate students.

## Searching

We'll begin by searching for "distance":

```{r}
ipeds_dict("distance")
```

This yields quite a few results (we're only shown 10 of 70). However, looking
through the list, we find that _efdeexc_ represents, "Students enrolled
exclusively in distance education courses" and covers 2011 to 2022. We'll use
this variable.

```{r}
df <- ipeds_init() |>
  ipeds_select(efdeexc) |>
  ipeds_year(2022) |> 
  ipeds_get() |>
  as_tibble()
```
```{r}
df
```

Looking at our output, we find an issue: _unitid_ values are not unique to each
row. Instead, each _unitid_ has multiple values of _efdeexc_. Looking back at
the output from `ipeds_dict()`, we can see that the file from which these values
come, `EF2022A_DIST`, is long. That means we need to find the variable that
uniquely identifies the values in _efdeexc_.

There are two ways to do this:

1. Search the filename with the `ripeds` dictionary and look through the list of
   possible variables: `ipeds_dict("EF2022A_DIST", search_col = "filename",
   limit = Inf)`
1. Download the dictionary file associated with the complete data file provided
   by IPEDS: `ipeds_download_to_disk(".", "EF2022A_DIST", type = "dictionary")`
   
The advantage of the first is that you don't have to download extra files (which
will unzip into either a Microsoft Excel workbook or HTML file). The benefit of
the second is that IPEDS dictionary contain more information than is presented
by `ipeds_dict()`.

## Check complete file dictionary file and filter

For this example, we'll download the IPEDS dictionary file.

```r
ipeds_download_to_disk(".", "EF2022A_DIST", type = "dictionary")
```

Looking in the **_Frequencies_** tab of the unzipped file, we can see that
_efdelev_ values map to different student levels and combination of levels. To
select only undergraduates, we need to set `efdelev == 2`.

| varnumber | varname | codevalue | valuelabel                                      | frequency | percent |
|:---------:|:--------|----------:|:------------------------------------------------|----------:|--------:|
| 24816     | EFDELEV |         1 | All students total                              |     5,978 |   26.88 |
| 24816     | EFDELEV |         2 | Undergraduate total                             |     5,706 |   25.66 |
| 24816     | EFDELEV |         3 | Undergraduate, degree/certificate-seeking total |     5,703 |   25.64 |
| 24816     | EFDELEV |        11 | Undergraduate, non-degree/certificate-seeking   |     2,803 |   12.60 |
| 24816     | EFDELEV |        12 | Graduate                                        |     2,049 |    9.21 |

Note that `ripeds` converts all input to lowercase, so choosing either `EFDELEV`
or `efdelev` will work. Adding the filter, we successfully return the number of
exclusively distance education undergraduate students in 2022.

```{r}
df <- ipeds_init() |>
  ipeds_select(efdeexc) |>
  ipeds_filter(efdelev == 2) |> 
  ipeds_year(2022) |> 
  ipeds_get() |>
  as_tibble()
```

```{r}
df
```

